\LaTeXe\ guide for the Journal of Functional Programming

### 2019-10-13 papers/jfp/1-intro.tex
Start working on a revision

\begin{itemize}
\item Build systems vary on many axes, including: static vs dynamic
  dependencies; local vs cloud; deterministic vs non-deterministic build tasks;
  support for early cutoff; self-tracking build systems; and the type of
  persistent build information. In~\S\ref{sec-background} we identify some of these key
  properties, illustrated by four carefully-chosen build systems.

@@

In short, instead of seeing build systems as unrelated points in space, we now
see them as locations in a connected landscape, leading to a better
understanding of what they do and how they compare, and making it easier to
explore other points in the landscape.

We discuss engineering aspects in \S\ref{sec-engineering}, and related
work in \S\ref{sec-related}. This paper is an extended version of an earlier
conference paper~\cite{mokhov2018buildsystems}.

% Papers about ``frameworks'' are often fuzzy.  This one is not: all our
% abstractions are defined in Haskell, and we have (freely-available)
% executable models of all the build systems we describe.

>>>>

\begin{itemize}
\item Build systems vary on many axes, including: static vs dynamic
  dependencies; local vs cloud; deterministic vs non-deterministic build tasks;
  early cutoff; self-tracking build systems; and the type of persistently stored
  build information. In~\S\ref{sec-background} we identify some of these key
  properties, illustrated by four carefully-chosen build systems.

@@

In short, instead of seeing build systems as unrelated points in space, we now
see them as locations in a connected landscape, leading to a better
understanding of what they do and how they compare, and making it easier to
explore other points in the landscape. While we steer clear of many engineering
aspects of real build systems, in~\S\ref{sec-engineering} we show how they can
be incorporated into the presented abstractions. The related work is discussed
in~\S\ref{sec-related}.

This paper is an extended version of an earlier conference
paper~\cite{mokhov2018buildsystems}. The key changes compared to the earlier
version are: (i) we added further clarifications and examples
into~\S\ref{sec-abstractions}, in particular,~\S\ref{sec-task-examples} is
entirely new; (ii)~\S\ref{sec-scheduler} and~\S\ref{sec-rebuilder} are based on
the material from the conference paper but have been substantially expanded to
include further details and examples, as well as completely new material such
as~\S\ref{sec-step-traces}; (iii)~\S\ref{sec-experience} is completely new; (iv)
\S\ref{sec-failures} and \S\S\ref{sec-polymorphism}-\ref{sec-tracking-aspects}
are almost entirely new, and \S\ref{sec-non-determinism} has been revised.
The new material focuses on our experience and various important practical
considerations, hence justifying the "and Practice" part of the paper title.

### 2019-10-13 papers/jfp/3-abstractions.tex
Start working on a revision

\begin{figure}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Abstract store containing a key/value map and persistent build information
data Store i k v -- i = info, k = key, v = value
initialise :: i -> (k -> v) -> Store i k v
getInfo    :: Store i k v -> i
putInfo    :: i -> Store i k v -> Store i k v
getValue   :: k -> Store i k v -> v
putValue   :: Eq k => k -> v -> Store i k v -> Store i k v
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
data Hash v -- a compact summary of a value with a fast equality check
hash    :: Hashable v => v -> Hash v
getHash :: Hashable v => k -> Store i k v -> Hash v
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build tasks (see §3.2)
newtype Task  c k v = Task { run :: @\std{forall}@ f. c f => (k -> f v) -> f v }
type    Tasks c k v = k -> Maybe (Task c k v)
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build system (see §3.3)
type Build c i k v = Tasks c k v -> k -> Store i k v -> Store i k v
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build system components: a scheduler (see §4) and a rebuilder (see §5)
type Scheduler c i ir k v = Rebuilder c ir k v -> Build c i k v
type Rebuilder c   ir k v = k -> v -> Task c k v -> Task (MonadState ir) k v
\end{minted}
% \vspace{-3.5mm}
\caption{Type signatures of key build systems abstractions.}\label{fig-types}
% \vspace{-4.5mm}
\end{figure}

@@

Our first main abstraction is for \emph{task descriptions}:
\begin{minted}[xleftmargin=5pt]{haskell}
newtype Task  c k v = Task@\,@{@\,\blk{run}\,@::@\,@@\std{forall}@ f. c f => (k -> f v) -> f v@\,@}
type    Tasks c k v = k -> Maybe (Task c k v)
\end{minted}

>>>>

\begin{figure}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Abstract store containing a key/value map and persistent build information
data Store i k v -- i = info, k = key, v = value
initialise :: i -> (k -> v) -> Store i k v
getInfo    :: Store i k v -> i
putInfo    :: i -> Store i k v -> Store i k v
getValue   :: k -> Store i k v -> v
putValue   :: Eq k => k -> v -> Store i k v -> Store i k v
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
data Hash v -- a compact summary of a value with a fast equality check
hash    :: Hashable v => v -> Hash v
getHash :: Hashable v => k -> Store i k v -> Hash v
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build tasks (see §3.2)
data Task  c k v = Task (@\std{forall}@ f. c f => (k -> f v) -> f v)
type Tasks c k v = k -> Maybe (Task c k v)
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
run :: c f => Task c k v -> (k -> f v) -> f v
run (Task task) = task
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build system (see §3.3)
type Build c i k v = Tasks c k v -> k -> Store i k v -> Store i k v
\end{minted}
% \vspace{-1mm}
\begin{minted}[fontsize=\small]{haskell}
-- Build system components: a scheduler (see §4) and a rebuilder (see §5)
type Scheduler c i ir k v = Rebuilder c ir k v -> Build c i k v
type Rebuilder c   ir k v = k -> v -> Task c k v -> Task (MonadState ir) k v
\end{minted}
% \vspace{-3.5mm}
\caption{Type signatures of key build systems abstractions.}\label{fig-types}
% \vspace{-4.5mm}
\end{figure}

@@

Our first main abstraction is for \emph{task descriptions}:
\begin{minted}[xleftmargin=10pt]{haskell}
data Task  c k v = Task (@\std{forall}@ f. c f => (k -> f v) -> f v)
type Tasks c k v = k -> Maybe (Task c k v)
\end{minted}
\begin{minted}[xleftmargin=10pt]{haskell}
run :: c f => Task c k v -> (k -> f v) -> f v
run (Task task) = task
\end{minted}
